On the adhesivity properties of equivalence graphs\thanks{???????}

### 2025-02-10 calco/main.tex
cose

\usepackage{stmaryrd}
\usepackage{bbding}
%\Envelope
\newcommand{\cat}[1]{\ensuremath{\mathbf{#1}}}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont} %simboli matematici
\usepackage[all, cmtip]{xy}

@@

\usepackage{hyperref, cleveref}
\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}

@@

Finally, another  useful properties of $\mathcal{M}$-adhesive category are the following.
\begin{proposition}\label{prop:regmono}
	If $\X$ is $\mathcal{M}$-adhesive then the following are true:
	\begin{enumerate}
		\item every $\mathcal{M}$-pushout square is also a pullback;
		\item every arrow in $\mathcal{M}$ is a regular mono.
	\end{enumerate}
\end{proposition}
\begin{proof}
	\begin{enumerate}
		\item  Consider the following cube in which the bottom face is an $\mathcal{M}$-pushout.
		\[\xymatrix@C=15pt@R=9pt{&A\ar[dd]|\hole_(.65){\id{A}}\ar[rr]^{g} \ar[dl]_{\id{A}} && B \ar[dd]^{\id{B}} \ar[dl]_(.6){\id{B}} \\ A  \ar@{>->}[dd]_{m}\ar[rr]^(.65){g} & & B \ar@{>->}[dd]_(.3){n}\\&A\ar[rr]|\hole^(.65){g} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n} \\C \ar[rr]_{f} & & D}\]
		By construction the top face of the cube is a pushout and the back one a pullback. The left face is a pullback because $m$ is mono, thus the Van Kampen property yields that the front and the right faces are pullbacks too and the thesis follows.
		\item Let $m\colon X\mto Y$ be an arrow in $\mathcal{M}$, we can then take its pushout along itself, which, by the previous point, is also a pullback.
		\[\xymatrix{X \ar@{>->}[r]^{m} \ar@{>->}[d]_{m}& Y \ar@{>->}[d]^{h}\\ Y \ar@{>->}[r]^{k} & Z}\]
		It is now immediate to see that $m$ is the equalizer of $h$ and $k$.  
	\end{enumerate}
\end{proof}

@@

\begin{example}\label{ex:kernel_pairs_in_Set}
    In $\Set$, a kernel pair for a function $f: A\to B$ is the set
    \[
        K_f=\{(x, y) \in A \times A \mid f(x) = f(y)\}
    \]
    together with the canonical projection on the first and the second component of the pairs.
\end{example}

\begin{proposition}\label{prop:pairng_of_kernel_pairs_mono}
    Let $(K, p_1, p_2)$ be the kernel pair of $f: X \to Y$, and let $(X\times X, \pi_1, \pi_2)$ be the product of $X$ with itself. Then, the mediating arrow $\langle p_1, p_2\rangle : K \to X \times X$ is mono.
\end{proposition}

@@

\begin{proposition}\label{prop:kermono}
	An arrow $m\colon M\to X$ is mono if and only if $(M, MM, \id{M})$ is a kernel pair for it.
\end{proposition}

@@

\begin{corollary}\label{cor:kermono}
	$(K_f, \pi_f^1, \pi_f^2)$ is a kernel pair for $f\colon X\to Y$ if and only if, for each mono $m\colon Y\to Z$, $(K_f, \pi_f^1, \pi_f^2)$ is a kernel pair also for $m\circ f$.
\end{corollary}
\begin{proof}
    It is enough to see that, by \Cref{lemma:pullback_lemma} and \Cref{prop:kermono} the outer boundary of the following square is a pullback.
        \[\begin{tikzcd}[row sep=13 pt, column sep=13 pt]
    	{K_f} && X && X \\
    	\\
    	X && Y && Y \\
    	\\
    	X && Y && Z
    	\arrow["{\pi_f^2}", from=1-1, to=1-3]
    	\arrow["{\pi_f^2}"', from=1-1, to=3-1]
    	\arrow["{id{X}}", from=1-3, to=1-5]
    	\arrow["f", from=1-3, to=3-3]
    	\arrow["f", from=1-5, to=3-5]
    	\arrow["f", from=3-1, to=3-3]
    	\arrow["{id_X}"', from=3-1, to=5-1]
    	\arrow["{id_Y}", from=3-3, to=3-5]
    	\arrow["{id_Y}", from=3-3, to=5-3]
    	\arrow["m", from=3-5, to=5-5]
    	\arrow["f"', from=5-1, to=5-3]
    	\arrow["m"', from=5-3, to=5-5]
    \end{tikzcd}\]
    The leftward part of the statement follows by definition of monomorphism an \Cref{lemma:pullback_lemma}.   
\end{proof}

\begin{lemma}\label{lemma:kern_pairs_pres_pullbacks}
    Suppose the following situation, and that $f: X \to Y$ and $g: Z \to W$ have kernel pairs.
    \[
        \begin{tikzcd}
            X \ar[r, "h"] \ar[d, "f"swap] & Z \ar[d, "g"] \\
            Y \ar[r, "t"swap] & W
        \end{tikzcd}
    \]

@@

\begin{proposition}\label{prop:reg_epi_coeq}
    Let $e: X \to Y$ be a regular epimorphism in a category $\cat C$ with a kernel pair $(K, \pi_1, \pi_2)$. Then, $e$ is the coequalizer of $\pi_1$ and $\pi_2$.
\end{proposition}

@@

\begin{corollary}\label{cor:reg_epi_components_reg_epi_nat_trans}
    Let $\cat C$ be a category with pullbacks and $\phi : D \dot\to D'$ be a natural transformation between two functors $D, D': \cat{I \to C}$. If $\phi_i$ is a regular epi for every $i$, then $\phi$ is a regular epi.
\end{corollary}

@@

\begin{lemma}\label{lemma:nat_trans_reg_epi_canonical_arrow_reg_epi}
    Let $D, D': \cat{I \to C}$ be two diagrams, and let $((c_i)_{i \in \cat I}, C)$ and $((c_i')_{i\in \cat I}, C')$ be, respectively, the colimit of $D$ and $D'$. If $\cat C$ has all colimits, for diagrams of shape $\cat I$ and $\phi: D \dot\to D'$ is a natural transformation in which all components are regular epimorphisms, then, the arrow induced by $\phi$ from $C$ to $C'$ (\Cref{prop:nat_tran_induces_unique_arrow_between_colimits}) is a regular epimorphism too.
\end{lemma}

@@

@@

Let $\pro^\star$ be the functor sending $X$ to $X^\star\times X^\star$, then we can present $\hyp$ as a comma category.
\begin{proposition}\label{prop:com}
	$\hyp$ is isomorphic to $\comma{\id{\Set}}{\pro^\star}$
\end{proposition}
\begin{proof}
	 Define two functors $F\colon  \hyp \to \comma{\id{\Set}}{\pro^\star} $
	and $G\colon \comma{\id{\Set}}{\pro^\star}\to \hyp$	 as follows 
	\begin{align*}
		\functor[l]{(E_\mathcal{{G}}, V_\mathcal{{G}}, s_\mathcal{{G}}, t_\mathcal{{G}})}{(f,g)}{\left(E_\mathcal{{H}}, V_\mathcal{{H}}, s_\mathcal{{H}}, t_\mathcal{{H}}\right)}
		\functormapsto
		\rfunctor{\left(E_\mathcal{{G}}, V_\mathcal{{G}}, 	\left(s_\mathcal{{G}}, t_\mathcal{{G}}\right)\right)}{(f, g) }{\left(E_\mathcal{{H}}, V_\mathcal{{H}}, \left(s_\mathcal{{H}}, t_\mathcal{{H}}\right)\right)}
	\end{align*}
	 \begin{align*}
		\functor[l]{(E_\mathcal{{G}}, V_\mathcal{{G}}, p_{\mathcal{G}})}{(f,g)}{(E_\mathcal{{H}}, V_\mathcal{{H}}, p_{\mathcal{H}})}
		\functormapsto
		\rfunctor{(E_\mathcal{{G}}, V_\mathcal{{G}}, 	\pi_1\circ  p_{\mathcal{G}}, \pi_2\circ  p_{\mathcal{G}})}{(f, g) }{(E_\mathcal{{H}}, V_\mathcal{{H}}, 	\pi_1\circ  p_{\mathcal{H}}, \pi_2\circ  p_{\mathcal{H}})}
	\end{align*} 
	Now it is immediate to notice that they are one the inverse of the other.   
\end{proof}

\begin{corollary}\label{prop:hypadh}
	 $\hyp$ is an adhesive category.
\end{corollary}
\begin{proof}
	By hypothesis $(-)^\star$ preserves pullbacks, while $\pro$ is continuous by definition, thus the thesis follows from  \Cref{thm:slice-functors} and \Cref{prop:com}.   
\end{proof}

@@

\begin{corollary}\label{cor:monhyper}
A morphism $(h,k)$ is a mono in $\hyp$ if and only if both its components are injective functions.
\end{corollary}

@@

\begin{proposition}\label{cor:left}  The forgetful functor $U_{\X}$ which sends an hypergraph $\mathcal{G}$ to its object of nodes has a left adjoint $\Delta_{\X}$.
\end{proposition}

@@

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{example}
\begin{example}\label{exa_2} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

@@

Following \cite{bonchi2022string}, we can present $\hyp_\X$ as a category of functor over a suitable category.

@@

Given $F\colon \catname{H}\to \catname{Set}$ we can define
\[E_F:=\sum_{k,l\in \mathbb{N}}F(k,l)\]

@@

Let $\mathcal{G}_F$ be the resulting hypergraph. Now, every $\eta\colon F\rightarrow H$ in $\catname{Set}^{\catname{H}}$ has components $\eta_{k,l}\colon F(k,l)\to H(k,l)$, $\eta_{\bullet}\colon F(\bullet)\to H(\bullet)$. In particular we can define $\hat{\eta}\colon E_F\rightarrow E_H$ as $\sum_{k,l\in \mathbb{N}}\eta_{k,l}$. Now, on the one hand, we have
\[\begin{split}
	\pi^H_i\circ \eta^k_\bullet \circ  s^F_{k,l}&=\eta_\bullet \circ \pi^F_i \circ s^F_{k,l}\\&=\eta_\bullet \circ F(f_i)\\&=H(f_i)\circ \eta_{k,l}\\&=\pi^H_{k,i}\circ s^{H}_{k,l}\circ \eta_{k,l}
\end{split} \qquad \begin{split}
\pi^H_j\circ \eta^k_\bullet \circ  t^F_{k,l}&=\eta_\bullet \circ \pi^F_j \circ t^F_{k,l}\\&=\eta_\bullet \circ F(f_{k+j})\\&=H(f_{k+j})\circ \eta_{k,l}\\&=\pi^H_{l,j}\circ t^{H}_{k,l}\circ \eta_{k,l}
\end{split} \]
so that $\eta^k_\bullet \circ  s^F_{k,l}=s^{H}_{k,l}\circ \eta_{k,l}$ and $\eta^k_\bullet \circ  t^F_{k,l}= t^{H}_{k,l}\circ \eta_{k,l}$. On the other hand, these equalities give us the following:
\[\begin{split}
	\eta_{\bullet}^\star \circ s_{F}\circ a^F_{k,l}&=\eta_{\bullet}^\star \circ b^F_k\circ s^F_{k,l}\\&=b^H_k\circ \eta^k_{\bullet}\circ s^{F}_{k,l} \\&=b^H_{k}\circ s^H_{k,l}\circ \eta_{k,l}\\&=s_H\circ a^H_{k,l}\circ \eta_{k,l}\\&=	s_H\circ \hat{\eta}\circ a^F_{k,l}
\end{split} \qquad \begin{split}
	\eta_{\bullet}^\star \circ s_{F}\circ a^F_{k,l}&=\eta_{\bullet}^\star \circ b^F_k\circ t^F_{k,l}\\&=b^H_k\circ \eta^k_{\bullet}\circ t^{F}_{k,l} \\&=b^H_{k}\circ t^H_{k,l}\circ \eta_{k,l}\\&=t_H\circ a^H_{k,l}\circ \eta_{k,l}\\&=	t_H\circ \hat{\eta}\circ a^F_{k,l}
\end{split}\]
which, in turn, entail that  that the following squares commute
\[\xymatrix{E_F \ar[r]^{s_F} \ar@{.>}[d]_{\hat{\eta}}& F(\bullet)^\star \ar[d]^{\eta^\star_\bullet} & E_F \ar[r]^{t_F} \ar[d]_{\hat{\eta}} & F(\bullet)^\star \ar[d]^{\eta^\star_\bullet}\\ E_H \ar[r]_{s_H} & H(\bullet)^\star & E_H \ar[r]_{t_H}& H(\bullet)^\star}\]

This is the same as saying that $\eta$ induces a morphism $(\hat{\eta}, \eta_{\bullet})\colon \mathcal{G}_F\to \mathcal{G}_H$. It is now clear that sending $F$ to $\mathcal{G}_F$ and $\eta$ to $(\hat{\eta}, \eta_{\bullet})$ defines a faithful functor $\mathcal{G}_{-}\colon \Set^{\catname{H}}\to \hyp$.

It is now a a straightforward exercise to see that this functor it is actually an equivalence (see \cite{castelnovo2023thesis} for details), yielding the following result.

	\iffalse 
\begin{remark}\label{rem:utile}
Now, notice that in $\Set$ countable coproducts distributes over pullbacks, \Cref{rem:salv} we know that countably distributivity entails that, for every $k,l\in \mathbb{N}$, the arrow $b^F_k\times b^F_l$ is a coprojection, thus extensivity entails that the square is a pullback 
	\[\xymatrix@C=30pt{F(k,l) \ar[d]_{(s^F_{k,l}, t^F_{k,l})}  \ar[r]^{a^F_{k,l}}& E_F \ar[d]^{(s_F, t_F)} \\ F(\bullet)^k\times F(\bullet)^l \ar[r]_{b^F_k\times b^F_l} & F(\bullet)^\star\times F(\bullet)^\star}\]

	As in \Cref{rem:pb}, countably extensivity of $X$ guarantees that, for every $k\in \mathbb{N}$ following square is a pullback.	
	\[\xymatrix{F(\bullet)^k  \ar[r]^{b^F_k} \ar[d]_{!_{F(\bullet)^k}}& F(\bullet)^\star \ar[d]^{\lgh_{F(\bullet)}} \\ 1 \ar[r]_{o_k} & 1^\star}\]

	Now, notice that 
	\[\begin{split}
		\lgh_{F(\bullet)}\circ b^{F}_k\circ s^{F}_{k,l}&=o_{k}\circ !_{F(\bullet)^k}\circ s^F_{k,l}\\&=o_k\circ !_{F(k,l)}
	\end{split} \qquad \begin{split}
	\lgh_{F(\bullet)}\circ b^{F}_k\circ t^{F}_{k,l}&=o_{k}\circ !_{F(\bullet)^k}\circ t^F_{k,l}\\&=o_k\circ !_{F(k,l)}
	\end{split}\]
so that we have two diagrams:
	\[\xymatrix{F(k,l)\ar[r]^{s^F_{k,l}} \ar@{.>}[dr]^{u^1_{k,l}}  \ar@/_.5cm/[ddr]_{!_{F(k,l)}}&F(\bullet)^k \ar[dr]^{b^F_k}&& F(k,l) \ar@/_.5cm/[ddr]_{!_{F(k,l)}} \ar[r]^{s^F_{k,l}} \ar@{.>}[dr]^{u^2_{k,l}} &F(\bullet)^k \ar[dr]^{b^F_k}\\ & F(\bullet)_k  \ar[r]^{q_{k}} \ar[d]_{!_{F(\bullet)_k}}& F(\bullet)^\star \ar[d]^{\lgh_{F(\bullet)}} && F(\bullet)_k  \ar[r]^{q_{k} }\ar[d]_{!_{F(\bullet)_k}}& F(\bullet)^\star \ar[d]^{\lgh_{F(\bullet)}} \\ &  1 \ar[r]_{o_k} & 1^\star && 1 \ar[r]_{o_k} & 1^\star}\]

	Moreover we have
	\[\begin{split}
		s_F\circ a^F_{k,l}&=b^F_k\circ s^{F}_{k,l}\\&=
		q_{k}\circ u^1_{k,l}	\end{split} \qquad \begin{split}
		t_F\circ a^F_{k,l}&=b^F_k\circ t^{F}_{k,l}\\&=
		q_{k}\circ u^2_{k,l}	\end{split}  \]

	Therefore these two arrows, in turn, yield the dotted $v^1_{k,l}\colon F(k,l)\to F(\bullet)_k $ and $v^2_{k,l}\colon F(k,l)\to F(\bullet)_l$ as in the diagram below, where the right and bottom squares are pullbacks.
	\[\xymatrix@C=45pt{F(k,l)  \ar[dr]^{a^F_{k,l}} \ar@{.>}[r]^{v^2_{k,l}} \ar@{.>}[d]^{v^1_{k,l}} \ar@/^.8cm/[rr]^{u^2_{k,l}} \ar@/_.5cm/[dd]_{u^1_{k,l}}& P_{l} \ar[r]_{p_{l,1}} \ar[d]_{p_{l,2}}& F(\bullet)_{l} \ar[d]^{q_{l}}\\ P_k \ar[r]^{p_{k,2}} \ar[d]^{p_{k,1}}& E_F \ar[r]_{s_F}  \ar[d]^{t_F}& F(\bullet)^\star \\ F(\bullet)_k  \ar[r]_{q_{k}}& F(\bullet)^\star}\]

 By hypothesis, $\X$ is countably extensive, thus for every $k, l\in \mathbb{N}$, the following diagrams are pullbacks.
 \[\xymatrix{F(k,l) \ar[r]^{a^F_{k,l}} & E_F \ar[d]^{}\\ 
 F(\bullet)^k& F(\bullet)^\star}\]

 Let now $d_1\colon Z\to P_l$ and $d_2\colon Z\to P_k $ be two arrows such that
 \[p_{k,2}\circ d_2=p_{t,2}\circ d_1\]

 Then we have
 \begin{align*}
 t_{F}\circ 	p_{t_2}\circ d_1&= \end{align*}

 Showing, finally, that

\end{remark}
\fi

\begin{proposition}
	$\hyp_\X$ is equivalent to the category $\catname{\X}^{\catname{H}}$.
\end{proposition}

\iffalse 
\begin{proof}	
We already know that $\mathcal{G}_{-}$ is faithful, we still have to show that it is full and essentially surjective.
	\begin{itemize}
		\item For fullness, let $(f,g)\colon \mathcal{G}_F\to \mathcal{G}_{H}$ be a morphism of $\X$-hypergraphs. By \Cref{rem:utile}, for every $k,l\in \mathbb{N}$, the front square in diagram below is a pullback, yielding the dotted arrow $\eta_{k,l}\colon F(k,l)\to H(k,l)$.
		\[\xymatrix@C=10pt{F(k,l)   \ar[dd]_{(s^F_{k,l}, t^F_{k,l})}\ar[rr]^-{a^F_{k,l}} \ar@{.>}[dr]_{\eta_{k,l }}&& E_F \ar[dd]|\hole^(.7){(s_F, t_F)}  \ar[dr]^{f}\\ & H(k,l) \ar[dd]_(.3){(s^H_{k,l}, t^H_{k,l})}\ar[rr]^(.4){a^H_{k,l}} &&  E_H \ar[dd]^{(s_H, t_H)}\\F(\bullet)^k\times F(\bullet)^l \ar[dr]_{g^k\times g^l} \ar[rr]|\hole^(.65){b^F_{k,l}\times b^F_{k,l}}& &F(\bullet)^\star \times F(\bullet)^\star  \ar[dr]^{g^\star\times g^\star}\\&H(\bullet)^k\times H(\bullet)^l \ar[rr]_{b^H_k\times b^H_l} && H(\bullet)^\star \times H(\bullet)^\star  }\]

		As $\eta_{\bullet}\colon F(\bullet)\to H(\bullet)$ we can simply take  $g$.  Then  the collection of all the $\eta_{k,l}$ and of $\eta_\bullet$ defines a natural transformation $\eta\colon F\to H$. Indeed, given an arrow $f_i\colon(k,l)\to \bullet$ then we have :	
		\begin{align*}
			\eta_\bullet \circ F(f_i)&=g\circ F(f_i)
			\\&= \begin{cases}
			g\circ \pi^F_i\circ s^F_{k,l} & i <k\\
			g\circ \pi^F_i\circ t^F_{k,l} & \text{otherwise}
			\end{cases}
			\\&=
			\begin{cases}
		\pi^H_i\circ g^k\circ s^F_{k,l} & i <k\\
		\pi^H_i\circ g^k\circ t^F_{k,l} 	& \text{otherwise}
			\end{cases}\\&=
		\begin{cases}
	\pi^H_i\circ s^H_{k,l} \circ \eta_{k,l}& i <k\\
	\pi^H_i\circ t^H_{k,l} \circ \eta_{k,l}	& \text{otherwise}
	\end{cases}\\&=H(i)\circ \eta_{k,l}
		\end{align*}
		Finally, by construction it is clear that $(\hat{\eta}, \eta_{\bullet})=(f,g)$.

		\item  Let us turn to essential surjectivity. Given an hypergraph $\mathcal{G}=(E_\mathcal{G}, V_\mathcal{G}, s_\mathcal{G}, t_\mathcal{G})$ we can define $F_{\mathcal{G}}(\bullet):=V_{\mathcal{G}}$ and $F_{\mathcal{G}}(k,l)$ by the following pullback. Notice that, by \Cref{rem:salv} and countably extensivity, $(E_{\mathcal{{G}}}, \{a_{k,l}\}_{k,l\in \mathbb{N}})$ is a coproduct.
		\[\xymatrix@C=30pt{F_\mathcal{G}(k,l) \ar[d]_{p_{k,l}}  \ar[r]^{a_{k,l}}& E_{\mathcal{G}} \ar[d]^{(s_F, t_F)} \\ V_{\mathcal{G}}^k\times V_{\mathcal{G}}^l \ar[r]_{b_k\times b_l} & V_{\mathcal{G}}^\star\times V_{\mathcal{G}}^\star}\]

		Let $q_{i;k,l}\colon V_\mathcal{G}^k\times V_{\mathcal{G}}^l\to V_\mathcal{G} $ be the $i^{\text{th}}$ projection, then for every morphism $f_i\colon (k,l)\to \bullet$ in $\catname{H}$ we can define $F_{\mathcal{G}}(f_i)$ as $q_{i; k,l}\circ p_{k,l}$. $F_{\mathcal{G}}$ so defined is a functor $\catname{H}\to \catname{Set}$.

		Notice that $p_{k,l}=(s^{F_{\mathcal{G}}}_{k,l}, t^{F_{\mathcal{G}}}_{k,l})$. To see this, it is enough to compose with the projections:
		\begin{align*}
			q_{i;k,l}\circ p_{k,l}&=F_{\mathcal{G}}(f_i)\\&=\begin{cases}
				\pi^{F_{\mathcal{G}}}_{k, i} \circ F_{\mathcal{G}}(f_i) & i<k\\
				\pi^{F_{\mathcal{G}}}_{l, i} \circ F_{\mathcal{G}}(f_i) &\text{otherwise}
			\end{cases}\\&=\begin{cases}
		s^{F_{\mathcal{G}}}_{k,l} & i<k\\
		t^{F_{\mathcal{G}}}_{k,l} &\text{otherwise}
			\end{cases}\\&=q_{i;k,l}\circ (s^{F_{\mathcal{G}}}_{k,l}, t^{F_{\mathcal{G}}}_{k,l})
		\end{align*}
		In particular, this entails that 
		\[\begin{split}
	s^{F_{\mathcal{G}}}\circ a^{F_{\mathcal{G}}}_{k,l}&=b^{F_{\mathcal{G}}}_k \circ s^{F_{\mathcal{G}}}_{k,l}\\&= b_k \circ s^{F_{\mathcal{G}}}_{k,l}\\&=s_F\circ a_{k,l}
		\end{split} \qquad \begin{split}
	t^{F_{\mathcal{G}}}\circ a^{F_{\mathcal{G}}}_{k,l}&=b^{F_{\mathcal{G}}}_l \circ t^{F_{\mathcal{G}}}_{k,l}\\&= b_l \circ s^{F_{\mathcal{G}}}_{k,l}\\&=t_F\circ a_{k,l}
		\end{split}\]

>>>>

\usepackage{stmaryrd}
\usepackage{bbding}
%\Envelope
\newcommand{\cat}[1]{\ensuremath{\mathbf{#1}}}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
\usepackage{hyperref, cleveref} %simboli matematici
\usepackage[all, cmtip]{xy}

@@

\usepackage{tikz-cd}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}

@@

Finally, another  useful properties of $\mathcal{M}$-adhesive category are the following.
\begin{proposition}\label[proposition]{prop:regmono}
	If $\X$ is $\mathcal{M}$-adhesive then the following are true:
	\begin{enumerate}
		\item every $\mathcal{M}$-pushout square is also a pullback;
		\item every arrow in $\mathcal{M}$ is a regular mono.
	\end{enumerate}
\end{proposition}
\begin{proof}
	\begin{enumerate}
		\item  Consider the following cube in which the bottom face is an $\mathcal{M}$-pushout.
		\[\xymatrix@C=15pt@R=9pt{&A\ar[dd]|\hole_(.65){\id{A}}\ar[rr]^{g} \ar[dl]_{\id{A}} && B \ar[dd]^{\id{B}} \ar[dl]_(.6){\id{B}} \\ A  \ar@{>->}[dd]_{m}\ar[rr]^(.65){g} & & B \ar@{>->}[dd]_(.3){n}\\&A\ar[rr]|\hole^(.65){g} \ar@{>->}[dl]^{m} && B \ar@{>->}[dl]^{n} \\C \ar[rr]_{f} & & D}\]
		By construction the top face of the cube is a pushout and the back one a pullback. The left face is a pullback because $m$ is mono, thus the Van Kampen property yields that the front and the right faces are pullbacks too and the thesis follows.
		\item Let $m\colon X\mto Y$ be an arrow in $\mathcal{M}$, we can then take its pushout along itself, which, by the previous point, is also a pullback.
		\[\xymatrix{X \ar@{>->}[r]^{m} \ar@{>->}[d]_{m}& Y \ar@{>->}[d]^{h}\\ Y \ar@{>->}[r]^{k} & Z}\]
		It is now immediate to see that $m$ is the equalizer of $h$ and $k$.  
	\end{enumerate}
\end{proof}

@@

\begin{example}\label[example]{ex:kernel_pairs_in_Set}
    In $\Set$, a kernel pair for a function $f: A\to B$ is the set
    \[
        K_f=\{(x, y) \in A \times A \mid f(x) = f(y)\}
    \]
    together with the canonical projection on the first and the second component of the pairs.
\end{example}

\begin{proposition}\label[proposition]{prop:pairng_of_kernel_pairs_mono}
    Let $(K, p_1, p_2)$ be the kernel pair of $f: X \to Y$, and let $(X\times X, \pi_1, \pi_2)$ be the product of $X$ with itself. Then, the mediating arrow $\langle p_1, p_2\rangle : K \to X \times X$ is mono.
\end{proposition}

@@

\begin{proposition}\label[proposition]{prop:kermono}
	An arrow $m\colon M\to X$ is mono if and only if $(M, MM, \id{M})$ is a kernel pair for it.
\end{proposition}

@@

\begin{corollary}\label[corollary]{cor:kermono}
	$(K_f, \pi_f^1, \pi_f^2)$ is a kernel pair for $f\colon X\to Y$ if and only if, for each mono $m\colon Y\to Z$, $(K_f, \pi_f^1, \pi_f^2)$ is a kernel pair also for $m\circ f$.
\end{corollary}
\begin{proof}
    It is enough to see that, by \Cref{lemma:pullback_lemma} and \Cref{prop:kermono} the outer boundary of the following square is a pullback.
        \[\begin{tikzcd}[row sep=13 pt, column sep=13 pt]
    	{K_f} && X && X \\
    	\\
    	X && Y && Y \\
    	\\
    	X && Y && Z
    	\arrow["{\pi_f^2}", from=1-1, to=1-3]
    	\arrow["{\pi_f^2}"', from=1-1, to=3-1]
    	\arrow["{id{X}}", from=1-3, to=1-5]
    	\arrow["f", from=1-3, to=3-3]
    	\arrow["f", from=1-5, to=3-5]
    	\arrow["f", from=3-1, to=3-3]
    	\arrow["{id_X}"', from=3-1, to=5-1]
    	\arrow["{id_Y}", from=3-3, to=3-5]
    	\arrow["{id_Y}", from=3-3, to=5-3]
    	\arrow["m", from=3-5, to=5-5]
    	\arrow["f"', from=5-1, to=5-3]
    	\arrow["m"', from=5-3, to=5-5]
    \end{tikzcd}\]
    The leftward part of the statement follows by definition of monomorphism an \Cref{lemma:pullback_lemma}.   
\end{proof}

\begin{lemma}\label[lemma]{lemma:kern_pairs_pres_pullbacks}
    Suppose the following situation, and that $f: X \to Y$ and $g: Z \to W$ have kernel pairs.
    \[
        \begin{tikzcd}
            X \ar[r, "h"] \ar[d, "f"swap] & Z \ar[d, "g"] \\
            Y \ar[r, "t"swap] & W
        \end{tikzcd}
    \]

@@

\begin{proposition}\label[proposition]{prop:reg_epi_coeq}
    Let $e: X \to Y$ be a regular epimorphism in a category $\cat C$ with a kernel pair $(K, \pi_1, \pi_2)$. Then, $e$ is the coequalizer of $\pi_1$ and $\pi_2$.
\end{proposition}

@@

\begin{corollary}\label[corollary]{cor:reg_epi_components_reg_epi_nat_trans}
    Let $\cat C$ be a category with pullbacks and $\phi : D \dot\to D'$ be a natural transformation between two functors $D, D': \cat{I \to C}$. If $\phi_i$ is a regular epi for every $i$, then $\phi$ is a regular epi.
\end{corollary}

@@

\begin{lemma}\label[lemma]{lemma:nat_trans_reg_epi_canonical_arrow_reg_epi}
    Let $D, D': \cat{I \to C}$ be two diagrams, and let $((c_i)_{i \in \cat I}, C)$ and $((c_i')_{i\in \cat I}, C')$ be, respectively, the colimit of $D$ and $D'$. If $\cat C$ has all colimits, for diagrams of shape $\cat I$ and $\phi: D \dot\to D'$ is a natural transformation in which all components are regular epimorphisms, then, the arrow induced by $\phi$ from $C$ to $C'$ (\Cref{prop:nat_tran_induces_unique_arrow_between_colimits}) is a regular epimorphism too.
\end{lemma}

@@

\begin{remark}\label[remark]{rem:length}Notice that, $1^\star$ can be canonically identified with $\mathbb{N}$, thus for every set $X$, the arrow $!_\X\colon X\to 1$ induces a \emph{length function} $!^\star_X\colon X^\star \to \mathbb{N}$ which simply sends a word length.
\end{remark}

@@

Let $\pro^\star$ be the functor sending $X$ to $X^\star\times X^\star$, then we can present $\hyp$ as a comma category.
\begin{proposition}\label[proposition]{prop:com}
	$\hyp$ is isomorphic to $\comma{\id{\Set}}{\pro^\star}$
\end{proposition}

\begin{corollary}\label[corollary]{prop:hypadh}
	 $\hyp$ is an adhesive category.
\end{corollary}
\begin{proof}
	By hypothesis $(-)^\star$ preserves pullbacks, while $\pro$ is continuous by definition, thus the thesis follows from  \Cref{thm:slice-functors} and \Cref{prop:com}.   
\end{proof}

@@

\begin{corollary}\label[corollary]{cor:monhyper}
A morphism $(h,k)$ is a mono in $\hyp$ if and only if both its components are injective functions.
\end{corollary}

@@

\begin{proposition}\label[proposition]{cor:left}  The forgetful functor $U_{\hyp}$ which sends an hypergraph $\mathcal{G}$ to its object of nodes has a left adjoint $\Delta_{\hyp}$.
\end{proposition}

@@

			\draw[->-=.5](1.75,-0.9)..controls(1.8,-1.3)and(2.5,-1.5)..(D)node[pos=0.5, below,font=\fontsize{7}{0}\selectfont]{$2$};
			\draw[->-=.5] (1.75,-0.6)..controls(1.8,-0.25)and(2.5,0)..(C) node[pos=0.5, above,font=\fontsize{7}{0}\selectfont]{$1$};
		\end{tikzpicture}
	\end{center}
\end{example}
\begin{example}\label[example]{exa_2} Let $V_{\mathcal{G}}$ be as in the previous example and $E_{\mathcal{G}}=\{h_1, h_2, h_3\}$.	Then we define
	\[\begin{matrix}
		s_{\mathcal{G}}(h_1)\colon 0\to V_{\mathcal{G}}  & s_{\mathcal{G}}(h_1)=?_{V_\mathcal{G}} && s_{\mathcal{G}}(h_2)\colon 2\to V_{\mathcal{G}} & \begin{matrix}
			0 \mapsto v_1\\
			1\mapsto v_2
		\end{matrix}&& s_{\mathcal{G}}(h_3)\colon 2\to V_{\mathcal{G}} & \begin{matrix} 
			0 \mapsto v_1\\
			1\mapsto v_4	
		\end{matrix}\\
		t_{\mathcal{G}}(h_1)\colon 1\to V_{\mathcal{G}} & 
		0 \mapsto v_1 && 		t_{\mathcal{G}}(h_2)\colon 1\to V_{\mathcal{G}} & 0\mapsto v_3 &&   t_{\mathcal{G}}(h_3)\colon 1\to V_{\mathcal{G}} & 1\mapsto v_5
	\end{matrix}\]

@@

Following \cite{bonchi2022string}, we can present $\hyp$ as a category of functor over a suitable category.

@@

The idea now is that for every functor $F\colon \catname{H}\to \catname{Set}$ we can define
\[E_F:=\sum_{k,l\in \mathbb{N}}F(k,l)\]

@@

Let $\mathcal{G}_F$ be the resulting hypergraph. One can now show that  sending $F$ to $\mathcal{G}_F$ can be extended to an equivalence $\mathcal{G}_{-}\colon \Set^{\catname{H}}\to \hyp$ (see \cite{castelnovo2023thesis, castelnovo2022newcriterionmathcalmmathcalnadhesivity} for details).

\begin{proposition}
	$\hyp$ is equivalent to the category $\Set^{\catname{H}}$.
\end{proposition}
